---
title: "LLMs for Convergent Thinking"
author: "Harsh Kumar, Ewan Jordan, Ashton Anderson"
date: "`r Sys.Date()`"
always_allow_html: true
output:
  html_document:
    code_folding: hide
    fig_height: 5
    fig_width: 6
    number_sections: yes
    toc: yes
    toc_depth: 4
---


```{r setup, include = F}

library(tidyverse)
library(scales)
library(here)
library(knitr)
library(gridExtra)
library(patchwork)
library(reshape2)
library(lmerTest)
library(emmeans)
library(dplyr)
library(broom)
library(ggbeeswarm)
library(ggpubr)
library(rstatix)
library(broom)
```

# Set constants
```{r set-constants}
# set style
theme_set(theme_bw())

# set paths
figure_dir <- here("figures/")
data_dir <- here("data/")
```

# Load and clean data

```{r load-data}
hit_ids <- c("66dd0fe39bfb33794b562bc9") # fourth and final wave

# read participants data
participants <- hit_ids %>%
  map_df(~read_csv(here(data_dir, sprintf("participants_%s.csv", .))))

# read responses
responses <- hit_ids %>%
  map_df(~read_csv(here(data_dir, sprintf("responses_%s.csv", .))))

# read verbal fluency responses
fluency <- hit_ids %>%
  map_df(~read_csv(here(data_dir, sprintf("fluency_%s.csv", .))))

# relabel condition
condition_labels <- c("absent" = "None",
                          "generate" = "LLM Answer",
                          "coach" = "LLM Guidance")
participants <- participants %>%
  mutate(condition = factor(condition,
                                levels = names(condition_labels),
                                labels = condition_labels))
responses <- responses %>%
  mutate(condition = factor(condition,
                                levels = names(condition_labels),
                                labels = condition_labels))

# reorder phase as a factor with practice and then test
phase_labels <- c("practice" = "Exposure", "test" = "Test")
responses <- responses %>%
  mutate(phase = factor(phase,
                        levels = names(phase_labels),
                        labels = phase_labels))
```
There were `r nrow(participants)` people who completed the experiment in total.

## Remove anyone who looks as though they were spam answering the questions

Excluding data from participants who fail to correctly answer the “attention check” question, after attempting the 4 questions, where participants were asked to select “numbers with units” which they came across during the task. The correct answer to this question is “none of these” options, so participants who answered otherwise will be excluded from our analyses.

```{r filter}
# Filter participants based on the survey_screener column  
participants <- participants %>%  
  filter(q3 == "Agree")  
  
# Filter responses based on the worker_ids of the filtered participants  
responses <- responses %>%  
  filter(workerid %in% participants$workerid)
```

# Sanity checks

## Randomization check

Check whether randomization is balanced across conditions (p-value should be large, indicating that the data are likely under the null of random assignment).

```{r check-random-assignment}
counts <- participants %>%
  count(condition)

kable(counts)

chisq.test(counts$n)
```
```{r accuracy-by-condition-phase}
plot_data <- responses %>%
  group_by(workerid, phase, condition) %>%
  summarize(frac_correct = mean(is_correct)) %>%
  group_by(phase, condition) %>%
  summarize(mean_frac_correct = mean(frac_correct) * 100,  # Scale to 0-100
            se_frac_correct = (sd(frac_correct) / sqrt(n())) * 100)  # Scale SE to 0-100

ggplot(plot_data, aes(x = phase, y = mean_frac_correct, group = condition, color = condition)) +  # Map color to condition
  geom_point(aes(shape = condition), size = 6, position = position_dodge(width = 0.2)) +  # Increase size of the dots
  geom_line(aes(linetype = condition), size = 1.2, position = position_dodge(width = 0.2)) +  # Slightly thicker lines
  geom_errorbar(aes(ymin = mean_frac_correct - se_frac_correct, ymax = mean_frac_correct + se_frac_correct),
                size = 1.2, width = 0.2, position = position_dodge(width = 0.2)) +  # Increase thickness of error bars
  scale_shape_manual(values = c(15, 16, 17)) +  # Different shapes for each condition
  scale_color_manual(values = c("#7F7F7F", "#e57a77", "#3d65a5")) +  # Lighter gray for the first condition
  scale_linetype_manual(values = c("solid", "dashed", "dotted")) +  # Different line types
  scale_x_discrete(expand = c(0.15, 0)) +  # Reduce the gap between phases
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Y-axis percentage scale
  coord_cartesian(clip = "off") +  # Ensure elements near the border are not cut off
  labs(x = 'Phase', y = 'Accuracy (%)') +
  theme_minimal(base_size = 12) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),  # Add a frame
    panel.grid.major = element_blank(),  # Remove major grid lines
    panel.grid.minor = element_blank(),  # Remove minor grid lines
    axis.line = element_line(size = 0.5),  # Emphasize the axis lines
    legend.position = c(0.95, 0.95),  # Top-right corner inside the plot
    legend.justification = c("right", "top"),  # Align legend to the top right inside the plot
    legend.title = element_blank(),  # Remove legend title for simplicity
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),  # Semi-transparent background for the legend
    axis.text = element_text(size = 22),  # Increase font size for axis labels
    axis.title = element_text(size = 26),  # Increase font size for axis titles
    legend.text = element_text(size = 20)  # Increase font size for legend text
  )
```
```{r calculate-fluency-prepare-model-data}
# Count the number of words listed by each participant
fluency_count <- fluency %>%
  group_by(workerid) %>%
  summarize(verbal_fluency = n())  # Count the number of entries per worker

# Merge the fluency count with the main dataset
responses <- responses %>%
  left_join(fluency_count, by = "workerid") %>%
  mutate(verbal_fluency = ifelse(is.na(verbal_fluency), 0, verbal_fluency))  # Replace NA in verbal_fluency with 0

# Use only the test data for the model
model_data <- responses %>%
  filter(phase == "Test")
```

```{r ancova}
model <- aov(is_correct ~ verbal_fluency + condition, data = model_data)
summary(model)
```
```{r post-hoc-comparisons}
# Run Tukey's HSD for pairwise comparisons of the 'condition' factor
tukey_results <- TukeyHSD(model, "condition")

# Display Tukey's HSD results
print(tukey_results)
```

```{r self-perceived-creativity}
# Calculate the difference between How Creative?-2 and How Creative?-1
participants <- participants %>%
  mutate(creative_diff = `How Creative?-2` - `How Creative?-1`)

# Calculate mean and SEM for each condition
summary_data <- participants %>%
  group_by(condition) %>%
  summarize(
    mean_diff = mean(creative_diff),
    sem_diff = sd(creative_diff) / sqrt(n())
  )

plot_self_creativity <- ggplot() +
  geom_point(data = summary_data, aes(x = condition, y = mean_diff, shape = condition, color = condition), size = 6) +  
  geom_errorbar(data = summary_data, aes(x = condition, ymin = mean_diff - sem_diff, ymax = mean_diff + sem_diff, color = condition), 
                size = 1.7, width = 0.25) +  
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size = 0.5) +  # Add horizontal line at y=0
  scale_shape_manual(values = c(15, 16, 17)) +  
  scale_color_manual(values = c("#7F7F7F", "#e57a77", "#3d65a5")) +  
  labs(x = 'Condition', y = 'Change in Reported\nSelf-Creativity Rating') +
  ylim(-8, 2) +  # Set the y-axis limits
  theme_minimal(base_size = 12) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),  
    panel.grid.major = element_blank(),  
    panel.grid.minor = element_blank(),  
    axis.line = element_line(size = 0.5),  
    legend.position = "none",  
    axis.text.x = element_text(size = 18),  # Customize x-axis text font size
    axis.text.y = element_text(size = 18),  # Customize y-axis text font size
    axis.title.x = element_text(size = 26),  # Customize x-axis title font size
    axis.title.y = element_text(size = 22),  # Customize y-axis title font size  
    legend.text = element_text(size = 20)
  )

plot_self_creativity

```
```{r ai-use}
# Convert Likert scale responses to numeric values
participants <- participants %>%
  mutate(pre_ai_feel = recode(`Increased AI use makes you feel-1`,
                              "More concerned than excited" = -1,
                              "Equally excited and concerned" = 0,
                              "More excited than concerned" = 1),
         post_ai_feel = recode(q2,
                               "More concerned than excited" = -1,
                               "Equally excited and concerned" = 0,
                               "More excited than concerned" = 1))

# Calculate the difference between post- and pre-experiment feelings
participants <- participants %>%
  mutate(feel_diff = post_ai_feel - pre_ai_feel)

# Calculate mean and SEM for the difference by condition
summary_data_feel_diff <- participants %>%
  group_by(condition) %>%
  summarize(
    mean_diff = mean(feel_diff),
    sem_diff = sd(feel_diff) / sqrt(n())
  )

# Plotting the mean difference with SEM
plot_ai_use <- ggplot() +
  #geom_quasirandom(data = participants, aes(x = condition, y = feel_diff, shape = condition, color = condition), size = 2, alpha = 0.6) +
  geom_point(data = summary_data_feel_diff, aes(x = condition, y = mean_diff, color = condition, shape = condition), size = 6) +  # Larger points
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size = 0.5) +  # Add horizontal line at y=0
  geom_errorbar(data = summary_data_feel_diff, aes(x = condition, ymin = mean_diff - sem_diff, ymax = mean_diff + sem_diff, color = condition), 
                size = 1.2, width = 0.2) +  # Thicker error bars with horizontal tips
  scale_shape_manual(values = c(15, 16, 17)) +  # Keep consistent shapes across conditions
  scale_color_manual(values = c("#7F7F7F", "#e57a77", "#3d65a5")) +  # Color-blind friendly palette
  scale_y_continuous(breaks = c(-2, -1, 0, 1, 2), labels = c("-2 (More concerned)", "-1", "0", "1", "2 (More excited)")) +  # Adjust y-axis labels
  labs(x = 'Condition', y = 'Change in Feeling\nabout AI Use') +  # Use delta (Δ) notation
  ylim(-0.1, 0.1) +  # Set the y-axis limits
  theme_minimal(base_size = 12) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),  # Add a frame
    panel.grid.major = element_blank(),  # Remove major grid lines
    panel.grid.minor = element_blank(),  # Remove minor grid lines
    
    axis.line = element_line(size = 0.5),  # Emphasize the axis lines
    legend.position = "none", 
    legend.title = element_blank(),  # Remove legend title for simplicity
    
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),  # Semi-transparent background for the legend
    axis.text.x = element_text(size = 19),  # Customize x-axis text font size
    axis.text.y = element_text(size = 20),  # Customize y-axis text font size
    axis.title.x = element_text(size = 26),  # Customize x-axis title font size
    axis.title.y = element_text(size = 22),  # Customize y-axis title font size  
    legend.text = element_text(size = 20)
    
  )
  
plot_ai_use
ggsave(filename="feeling.png")
```

```{r perceived-difficulty}
plot_difficulty <- participants %>%
  mutate(difficulty = factor(q1, levels = c("Very easy", "Somewhat easy", "Somewhat difficult", "Very difficult"))) %>%
  group_by(condition) %>%
  summarize(mean_difficulty = mean(as.numeric(difficulty)),
            se = sd(as.numeric(difficulty)) / sqrt(n()), .groups = 'drop') %>%
  ggplot(aes(x = condition, y = mean_difficulty, color = condition, shape = condition)) +
  geom_point(, size = 6) +  # Increase the size of the points
  geom_errorbar(aes(ymin = mean_difficulty - se, ymax = mean_difficulty + se), 
                size = 1.2, width = 0.2) +  # Increase thickness and add horizontal tips
  scale_shape_manual(values = c(15, 16, 17)) +  # Keep consistent shapes across conditions
  scale_color_manual(values = c("#7F7F7F", "#e57a77", "#3d65a5")) +  # Color-blind friendly palette
  labs(x = 'Condition', y = 'Perceived Difficulty\nof Test Rounds') +
  theme_minimal(base_size = 12) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),  # Add a frame
    panel.grid.major = element_blank(),  # Remove major grid lines
    panel.grid.minor = element_blank(),  # Remove minor grid lines
    axis.line = element_line(size = 0.5),  # Emphasize the axis lines
    legend.position = "none",  # Remove the legend for simplicity
    axis.text.x = element_text(size = 19),  # Customize x-axis text font size
    axis.text.y = element_text(size = 20),  # Customize y-axis text font size
    axis.title.x = element_text(size = 26),  # Customize x-axis title font size
    axis.title.y = element_text(size = 22),  # Customize y-axis title font size  
    legend.text = element_text(size = 20)
  )

plot_difficulty
```


```{r perceived-helpfulness-2}
# Categorize responses as "Helpful" or "Not Helpful"
participants <- participants %>%
  mutate(helpful = ifelse(q4 %in% c("A little helpful", "Very helpful"), 1, 0))  # 1 for Helpful, 0 for Not Helpful

# Calculate mean percentage and SEM for helpfulness by condition
summary_data_helpfulness_pct <- participants %>%
  group_by(condition) %>%
  summarize(
    mean_helpful_pct = mean(helpful) * 100,  # Mean percentage
    sem_helpful_pct = sd(helpful) / sqrt(n()) * 100  # SEM percentage
  )

# Plotting the mean percentage of participants who found the practice round helpful with SEM
plot_helpfulness <- ggplot(summary_data_helpfulness_pct, aes(x = condition, y = mean_helpful_pct, color = condition)) +
  geom_point(aes(shape = condition), size = 6) +  # Point for mean percentage, increased size
  geom_errorbar(aes(ymin = mean_helpful_pct - sem_helpful_pct, ymax = mean_helpful_pct + sem_helpful_pct), 
                size = 1.2, width = 0.2) +  # Error bars with increased thickness and horizontal tips
  #geom_quasirandom(data = participants, aes(x = condition, y = helpful * 100, shape = condition), size = 2, alpha = 0.6) +  # Overlay raw values
  scale_shape_manual(values = c(15, 16, 17)) +  # Keep consistent shapes across conditions
  scale_color_manual(values = c("#7F7F7F", "#e57a77", "#3d65a5")) +  # Color-blind friendly palette
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Y-axis as percentage
  labs(x = 'Condition',y = '% of Participants Who Found\nExposure Rounds Helpful') +
  theme_minimal(base_size = 12) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),  # Add a frame
    panel.grid.major = element_blank(),  # Remove major grid lines
    panel.grid.minor = element_blank(),  # Remove minor grid lines
    axis.line = element_line(size = 0.5),  # Emphasize the axis lines
    legend.position = "none",  # Remove the legend for simplicity
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),  # Semi-transparent background for the legend
    axis.text.x = element_text(size = 19),  # Customize x-axis text font size
    axis.text.y = element_text(size = 20),  # Customize y-axis text font size
    axis.title.x = element_text(size = 26),  # Customize x-axis title font size
    axis.title.y = element_text(size = 22),  # Customize y-axis title font size  
    legend.text = element_text(size = 20)
  )

plot_helpfulness
```


# Save variables for .rtex document



# R Session information with package versions

Useful for reproducibility purposes

```{r session-info}

sessionInfo()

```